Notes on Perturbational Quantities

### 2008-07-30 docs/pert_notes/paper.tex
Here's my (current) best guess about how to handle predicting (rho h)prime to edges.  It mirrors the treatment for rhoprime.  But there is a difference, since we compute (rho h)_0 by averaging (rho h)^{(1)}, we do not need to deal with any etarhoh stuff to correct rhoh.

We predict $\rho'$ to edges using make\_edge\_scal in scalar\_advance.  The underbraced 
forcing term in (\ref{rhoprime equation}) is computed in modify\_scal\_force.

@@

We first compute $\rho_0^{(2)}$ from $\rho^n$ using advect\_base.  The $(2)$ in the 

@@

+ \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext}. \label{rhoh equation}

@@

Full temperature equation:

@@

These are working notes for what I'd like to do for the energy evolution.

@@

+ \frac{D_0p_0}{Dt} + \overline{\nabla\cdot\kappa\nabla T} + \overline{\rho H_{\rm nuc}} 
+ \overline{\rho H_{\rm ext}} \nonumber \\

@@

+ \psi + \overline{\nabla\cdot\kappa\nabla T} + \overline{\rho H_{\rm nuc}} 
+ \overline{\rho H_{\rm ext}}.\label{rhoh0 equation}

@@

We will predict both $(\rho h)_0$ and $(\rho h)'$ to edges using a Godunov method.  Then 
we will add the two answers together to get $\rho h$ at edges.  This is different from 
the density, in which we averaged an ``old'' and ``new'' base state density to get 
$\rho_0$ at edges.  By computing $\rho h$ at edges in this way, we avoid having to 
compute the complete evolution of $(\rho h)_0$, and therefore avoid having to write a 
correct\_base and diffuse\_base function for $\rho h$.\\

To simplify matters, the we have already called react\_state, so for now we will ignore 
the $\rho H_{\rm nuc}$ and $\rho H_{\rm ext}$ terms in the remaining derivations.  
Subtracting (\ref{rhoh0 equation}) from (\ref{rhoh equation}) and ignoring the 
aforementioned terms gives the perturbational enthalpy equation:
\begin{eqnarray}
\frac{\partial(\rho h)'}{\partial t} &=& -\nabla\cdot[(\rho h)'\Ub] 
- \nabla\cdot[(\rho h)_0\Ubt] + \wt\frac{\partial p_0}{\partial r} 
+ (\nabla\cdot\kappa\nabla T - \overline{\nabla\cdot\kappa\nabla T}) \nonumber \\
&=& -\Ub\cdot\nabla(\rho h)' \underbrace{- (\rho h)'\nabla\cdot\Ub 
- \nabla\cdot[(\rho h)_0\Ubt]} + \wt\frac{\partial p_0}{\partial r} 
+ (\nabla\cdot\kappa\nabla T 
- \overline{\nabla\cdot\kappa\nabla T})\label{rhohprime equation}
\end{eqnarray}
We use make\_edge\_scal to get $(\rho h)'$ at edges.  The function modify\_scal\_force 
computes the underbraced term in (\ref{rhohprime equation}).  The function mkrhohforce 
computes the $\wt$ term.\MarginPar{I need to write a function to handle the thermal 
diffusion terms.}
\subsubsection{Predicting $(\rho h)_0$ at edges}\label{Predicting rhoh0 at edges}
After the call to react\_state, we set $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$.  
We use the Godunov algorithm in make\_edge\_state\_1d to predict $(\rho h)_0$ at edges, 
omitting the terms added in react\_state, so we are essentially evolving an abbreviated 
version of (\ref{rhoh0 equation}):
\begin{eqnarray}
\frac{\partial(\rho h)_0}{\partial t} &=& -\nabla\cdot[(\rho h)_0 w_0\eb_r] + \psi 
+ \overline{\nabla\cdot\kappa\nabla T} \nonumber \\
&=& -w_0\frac{\partial(\rho h)_0}{\partial r} - (\rho h)_0\frac{\partial w_0}{\partial r} 
+ \psi + \overline{\nabla\cdot\kappa\nabla T}.\label{rhoh0 equation}
\end{eqnarray}
In advect\_base, the forcing term is set to the term proportional to 
$(\rho h)_0$.\MarginPar{We are currently omitting the $\psi$ and thermal conduction 
terms.  I think this is a bug.}
\subsubsection{Computing $\rho h$ at edges}
Simply add the results from Sections \ref{Predicting rhohprime at edges} and 
\ref{Predicting rhoh0 at edges}, using spatial interpolation to get $(\rho h)_0$ 
at non-radial edges.
\subsection{Predicting $h$ (enthalpy\_pred\_type = 2)}
\subsection{Predicting $T$ into $(\rho h)'$ (enthalpy\_pred\_type = 3)}
\subsection{Predicting $T$ into $h$ (enthalpy\_pred\_type = 4)}
\end{document}

>>>>

We define $\rho'^n = \rho^n - \rho_0^n$.  Then we predict $\rho'$ to edges using 
make\_edge\_scal in scalar\_advance and the underbraced term in (\ref{rhoprime equation}) 
as the forcing.

@@

We first compute $\rho_0^{(2)}$ from $\rho^n$ using advect\_base, which evolves 
(\ref{rho0 equation}) through $\Delta t$ in time.  The $(2)$ in the 

@@

+ \nabla\cdot\kappa\nabla T + \rho H_{\rm nuc} + \rho H_{\rm ext}.

@@

The full temperature equation is:

@@

These are working notes for how we are going to handle the energy evolution.  Due to the 
operator splitting, the call to react\_state has already been made.  Hence, the goal is to 
compute an edge state energy starting from $(\rho h)^{(1)}$ using an energy evolution 
equation that does not include terms already accounted for in react\_state
\frac{\partial(\rho h)}{\partial t} = -\Ub\cdot\nabla(\rho h) - \rho h\nabla\cdot\Ub 
+ \psi + \wt\frac{\partial p_0}{\partial r} + \nabla\cdot\kappa\nabla T. \label{rhoh equation}

@@

+ \frac{D_0p_0}{Dt} \nonumber \\

@@

+ \psi.\label{rhoh0 equation}

@@

Subtracting (\ref{rhoh0 equation}) from (\ref{rhoh equation}) gives the perturbational 
enthalpy equation:
\frac{\partial(\rho h)'}{\partial t} &=& -\nabla\cdot[(\rho h)'\Ub] 
- \nabla\cdot[(\rho h)_0\Ubt] + \wt\frac{\partial p_0}{\partial r} 
+ \nabla\cdot\kappa\nabla T\nonumber \\
&=& -\Ub\cdot\nabla(\rho h)' \underbrace{- (\rho h)'\nabla\cdot\Ub 
- \nabla\cdot[(\rho h)_0\Ubt] + \wt\frac{\partial p_0}{\partial r}
+ \nabla\cdot\kappa\nabla T}_{"(\rho h)' ~ \text{force}"} \label{rhohprime equation}
We define $(\rho h)_0^{(1)} = \overline{(\rho h)^{(1)}}$  and 
$(\rho h)'^{(1)} = (\rho h)^{(1)}-(\rho h)_0^{(1)}$.  Then we predict $(\rho h)'^{(1)}$ to 
edges using make\_edge\_scal in scalar\_advance and the underbraced term in 
(\ref{rhohprime equation}) as the 
forcing.
\subsubsection{Computing $\rho h$ at edges}
We first compute $(\rho h)_0^{(2)}$ from $(\rho h)_0^{(1)}$ using advect\_base, which evolves 
(\ref{rhoh equation}) through $\Delta t$ in time.  Then we follow an analogous procedure 
used in Section \ref{Computing rho at edges} to obtain $\rho h$ at edges.
\subsection{Predicting $h$ (enthalpy\_pred\_type = 2)}
\subsection{Predicting $T$ into $(\rho h)'$ (enthalpy\_pred\_type = 3)}
\subsection{Predicting $T$ into $h$ (enthalpy\_pred\_type = 4)}
\end{document}
